\section{The \GeoGen predicate language v1}
\label{app:property_reference}

\GeoGenBench's predicate library is the first published version of the \GeoGen predicate language.
Future instantiations of the methodology (\cref{sec:discussion}) will extend the vocabulary --- adding face-incidence for 3D solids, point-on-curve for coordinate plane, transformation-equivalence for symmetry constructions --- but the predicate-language design pattern (typed signatures, declarative semantics, tolerance-aware decision procedures, optional soft predicates for source-string checks that cannot reach into the rendered output) is intended to remain stable.
This appendix is the reference for v1.

For each predicate we give: (a)~the formal signature with positional argument types; (b)~the decision rule, including tolerance handling; (c)~the implementation reference (file:line in \texttt{util/tikz\_geometry.py} unless otherwise noted); and (d)~where applicable, soft-spot notes from \cref{ssec:ver_defn} and the predicate's verdict on an adversarial input.
\Cref{tab:predicates} in \cref{ssec:properties} summarises the v1 vocabulary at a glance.
All decisions use the production tolerance $\tau = 5\times 10^{-3}$ unless overridden by the per-check \texttt{tol} field in the scenario YAML.

In what follows we use $\mathbf{X}: \text{name} \to (x, y) \in \mathbb{R}^2$ for the coordinate map produced by the TikZ extractor, $|PQ|$ for the Euclidean distance between two coordinates, $\vec{PQ}$ for the vector $Q-P$, $\vec{u}\cdot\vec{v}$ for the dot product, and $\vec{u}\times\vec{v}$ for the scalar 2D cross product $u_x v_y - u_y v_x$.

\subsection{Symbolic predicates (decided exactly against \texorpdfstring{$\mathbf{X}$}{X})}

\subsubsection{\propname{right\_angle}}
\label{app:pred:right_angle}

\noindent\textbf{Signature.}\quad \texttt{right\_angle([a, vertex, c])} where each argument is a point name in $\mathbf{X}$.

\noindent\textbf{Semantics.}\quad The angle at \texttt{vertex} formed by rays to \texttt{a} and \texttt{c} is a right angle.

\noindent\textbf{Decision rule.}\quad Compute $\vec{u} = \vec{(\text{vertex})\,a}$ and $\vec{v} = \vec{(\text{vertex})\,c}$; return $|\vec{u}\cdot\vec{v}| \leq \tau$.
The dot-product test is preferred over an explicit angle computation because it has no $\arccos$ branch issues and is symmetric in the leg lengths.

\noindent\textbf{Reference.}\quad \texttt{util/tikz\_geometry.py:397}.

\noindent\textbf{Soft spots.}\quad None observed.
The dot-product threshold is tight enough that floating-point near-90$^\circ$ angles ($\sim$89.9$^\circ$) are correctly rejected at $\tau = 5\times 10^{-3}$ for the unit-scale coordinates produced by TikZ.

\subsubsection{\propname{midpoint}}
\label{app:pred:midpoint}

\noindent\textbf{Signature.}\quad \texttt{midpoint([m, a, b])}.

\noindent\textbf{Semantics.}\quad Point \texttt{m} is the midpoint of segment $\overline{ab}$.

\noindent\textbf{Decision rule.}\quad Return $\max(|m_x - (a_x{+}b_x)/2|,\ |m_y - (a_y{+}b_y)/2|) \leq \tau$.
We test componentwise on the expected midpoint coordinates rather than testing $|MA|=|MB|$, because the latter is satisfied by every point on the perpendicular bisector of $\overline{ab}$.

\noindent\textbf{Reference.}\quad \texttt{util/tikz\_geometry.py:406}.

\noindent\textbf{Soft spots.}\quad The earlier ``equal-distance'' implementation (now superseded) had a soft spot allowing $M$ anywhere on the perpendicular bisector --- see \cref{ssec:ver_defn}.
The current componentwise check eliminates that escape entirely.

\subsubsection{\propname{centroid}}
\label{app:pred:centroid}

\noindent\textbf{Signature.}\quad \texttt{centroid([g, a, b, c])}.

\noindent\textbf{Semantics.}\quad Point \texttt{g} is the centroid of $\triangle abc$, i.e., $g = (a + b + c)/3$.

\noindent\textbf{Decision rule.}\quad Return $\max(|g_x - (a_x{+}b_x{+}c_x)/3|,\ |g_y - (a_y{+}b_y{+}c_y)/3|) \leq \tau$.
The componentwise form mirrors \propname{midpoint}: testing $|GA|+|GB|+|GC|$ minimisation would admit any point on the locus where the sum-of-distances function is locally flat (which is in fact only the centroid, but the fixed-point form is cheaper and more obviously correct).

\noindent\textbf{Reference.}\quad \texttt{util/tikz\_geometry.py:418}.

\noindent\textbf{Soft spots.}\quad This is the canonical \emph{self-referential} predicate (\cref{ssec:self_referential_checks}): in the pilot, every observed implementation emits $G$ either through an IR \texttt{centroid} derivation (in \structured) or via $\backslash$\texttt{tkzInterLL} on two medians (in \rawcode), and both forms make the formula check tautological for any non-degenerate triangle.
The check is non-trivial only when a model writes literal numerical coordinates for $G$, which we did not observe in v1; we retain the predicate because the median-concurrency form (an alternative semantics that requires the three medians to be \emph{drawn}, not merely the centroid \emph{computed}) is a v1.1 candidate.

\subsubsection{\propname{collinear}}
\label{app:pred:collinear}

\noindent\textbf{Signature.}\quad \texttt{collinear([a, b, c])}.

\noindent\textbf{Semantics.}\quad The three points lie on a common line.

\noindent\textbf{Decision rule.}\quad Return $|\vec{ab}\times\vec{ac}| \leq \tau$.
The cross-product test is the signed area of $\triangle abc$ doubled; zero iff degenerate.

\noindent\textbf{Reference.}\quad \texttt{util/tikz\_geometry.py:417}.

\noindent\textbf{Soft spots.}\quad The check does not require $a \neq b$; three coincident points trivially pass.
Pair with \propname{distinct\_points} (IR-side) or rely on TikZ's failure to draw a zero-length segment if degeneracy matters for the prompt.

\subsubsection{\propname{equal\_lengths}}
\label{app:pred:equal_lengths}

\noindent\textbf{Signature.}\quad \texttt{equal\_lengths([[p1, p2], [p3, p4], \dots])} --- a list of $\geq 2$ point pairs, each defining a segment.

\noindent\textbf{Semantics.}\quad All listed segments have the same length.

\noindent\textbf{Decision rule.}\quad Compute $d_i = |p_{2i-1} p_{2i}|$ for each pair; return $\max_{i\geq 2} |d_i - d_1| \leq \tau$.
A list with a single pair returns \texttt{None} (cannot decide equality of one length).

\noindent\textbf{Reference.}\quad \texttt{util/tikz\_geometry.py:426}.

\noindent\textbf{Soft spots.}\quad Tolerance is absolute, not relative; for very long segments ($|PQ| \gg 1$) the absolute tolerance becomes effectively tighter, but every \GeoGenBench scenario keeps coordinates in the range $[-10, 10]$ so this is not observed.

\subsubsection{\propname{parallel}}
\label{app:pred:parallel}

\noindent\textbf{Signature.}\quad \texttt{parallel([[a, b], [c, d]])}.

\noindent\textbf{Semantics.}\quad Lines $\overleftrightarrow{ab}$ and $\overleftrightarrow{cd}$ are parallel.

\noindent\textbf{Decision rule.}\quad Return $|\vec{ab}\times\vec{cd}| \leq \tau$.

\noindent\textbf{Reference.}\quad \texttt{util/tikz\_geometry.py:438}.

\noindent\textbf{Soft spots.}\quad Coincident lines pass --- $\overleftrightarrow{ab}$ is parallel to itself.
For prompts where ``parallel'' implies ``parallel and distinct'', pair with \propname{point\_on\_line} = \texttt{False} for one endpoint of the second line against the first (see \cref{ssec:ver_defn}).

\subsubsection{\propname{perpendicular}}
\label{app:pred:perpendicular}

\noindent\textbf{Signature.}\quad \texttt{perpendicular([[a, b], [c, d]])}.

\noindent\textbf{Semantics.}\quad Lines $\overleftrightarrow{ab}$ and $\overleftrightarrow{cd}$ meet at right angles (or would if extended).

\noindent\textbf{Decision rule.}\quad Return $|\vec{ab}\cdot\vec{cd}| \leq \tau$.

\noindent\textbf{Reference.}\quad \texttt{util/tikz\_geometry.py:448}.

\noindent\textbf{Soft spots.}\quad None.
Note this does \emph{not} require the lines to actually intersect within the visible canvas; for that, pair with \propname{intersects}.

\subsubsection{\propname{point\_on\_line}}
\label{app:pred:point_on_line}

\noindent\textbf{Signature.}\quad \texttt{point\_on\_line([p, a, b])}.

\noindent\textbf{Semantics.}\quad Point \texttt{p} lies on the infinite line $\overleftrightarrow{ab}$.

\noindent\textbf{Decision rule.}\quad Return $|\vec{ab}\times\vec{ap}| \leq \tau$ (signed area of $\triangle abp$ is zero).

\noindent\textbf{Reference.}\quad \texttt{util/tikz\_geometry.py:456}.

\noindent\textbf{Soft spots.}\quad The check accepts $p$ on the line at \emph{any} parameter, including outside segment $\overline{ab}$.
For ``on segment'' semantics, use \propname{point\_on\_segment}.

\subsubsection{\propname{point\_on\_segment}}
\label{app:pred:point_on_segment}

\noindent\textbf{Signature.}\quad \texttt{point\_on\_segment([p, a, b])}.

\noindent\textbf{Semantics.}\quad Point \texttt{p} lies on segment $\overline{ab}$ (between $a$ and $b$, inclusive).

\noindent\textbf{Decision rule.}\quad Two-step:
\begin{enumerate}
\item Collinearity: $|\vec{ab}\times\vec{ap}| \leq \tau$. Return \false if not.
\item Between-ness: $\big||ap| + |pb| - |ab|\big| \leq \tau$ (triangle equality holds iff $p$ is on the segment).
\end{enumerate}

\noindent\textbf{Reference.}\quad \texttt{util/tikz\_geometry.py:464}.

\noindent\textbf{Soft spots.}\quad The pre-hardening implementation only ran step (1) --- this caused a measurable false-positive rate documented in \cref{ssec:ver_defn} and was the primary motivator for the hardening pass in \cref{ssec:ver_defn}.

\subsubsection{\propname{point\_on\_circle}}
\label{app:pred:point_on_circle}

\noindent\textbf{Signature.}\quad \texttt{point\_on\_circle([p, o, r])} where \texttt{o} is the centre and \texttt{r} is any reference point on the circle (typically the radius witness).

\noindent\textbf{Semantics.}\quad Point \texttt{p} lies on the circle centred at \texttt{o} with radius $|or|$.

\noindent\textbf{Decision rule.}\quad Return $\big||op| - |or|\big| \leq \tau$.

\noindent\textbf{Reference.}\quad \texttt{util/tikz\_geometry.py:476}.

\noindent\textbf{Soft spots.}\quad Requires the radius witness \texttt{r} to be a resolved coordinate; if \texttt{r} is itself a constructed point that the extractor cannot resolve, the check returns \texttt{None} (silent skip). Templates avoid this by passing a directly-defined point as \texttt{r}.

\subsubsection{\propname{tangent}}
\label{app:pred:tangent}

\noindent\textbf{Signature.}\quad \texttt{tangent([[l1, l2], o, t])} --- line through $l_1, l_2$, circle centre $o$, point of tangency $t$.

\noindent\textbf{Semantics.}\quad Line $\overleftrightarrow{l_1 l_2}$ is tangent to the circle centred at $o$, touching it at point $t$.

\noindent\textbf{Decision rule.}\quad Two-step:
\begin{enumerate}
\item $t$ on line: $|\vec{l_1 l_2}\times\vec{l_1 t}| \leq \tau$.
\item $\vec{ot} \perp \vec{l_1 l_2}$: $|\vec{l_1 l_2}\cdot\vec{ot}| \leq \tau$.
\end{enumerate}

\noindent\textbf{Reference.}\quad \texttt{util/tikz\_geometry.py:483}.

\noindent\textbf{Soft spots.}\quad Does not separately verify that $t$ lies on the circle (that $|ot|$ equals the circle's radius). For that, pair with \propname{point\_on\_circle} (\cref{ssec:ver_defn}).
Templates that emit \propname{tangent} also emit \propname{point\_on\_circle} on $t$ as a co-check.

\subsubsection{\propname{angle\_equal}}
\label{app:pred:angle_equal}

\noindent\textbf{Signature.}\quad \texttt{angle\_equal([[a1, b1, c1], [a2, b2, c2]])} --- two angle triples in (leg, vertex, leg) order.

\noindent\textbf{Semantics.}\quad $\angle a_1 b_1 c_1 = \angle a_2 b_2 c_2$ (unsigned, in $[0, \pi]$).

\noindent\textbf{Decision rule.}\quad For each triple, compute $\theta_i = \arccos\big(\vec{b_i a_i}\cdot\vec{b_i c_i} / (|\vec{b_i a_i}||\vec{b_i c_i}|)\big)$.
Return $|\theta_1 - \theta_2| \leq \tau$ (radians).

\noindent\textbf{Reference.}\quad \texttt{util/tikz\_geometry.py:499}.

\noindent\textbf{Soft spots.}\quad Compares unsigned angles, so a $30^\circ$ angle and a $-30^\circ$ angle (orientation-flipped) are equal under this predicate.
Templates needing oriented-angle equality must additionally constrain orientation (\propname{same\_side}, IR-side).

\subsubsection{\propname{angle\_bisector}}
\label{app:pred:angle_bisector}

\noindent\textbf{Signature.}\quad \texttt{angle\_bisector([d, a, b, c])}.

\noindent\textbf{Semantics.}\quad Ray $\vec{ad}$ bisects $\angle bac$.

\noindent\textbf{Decision rule.}\quad Compute $\theta_1 = \angle bad$ and $\theta_2 = \angle dac$ via the same $\arccos$ formula as \propname{angle\_equal}; return $|\theta_1 - \theta_2| \leq \tau$.

\noindent\textbf{Reference.}\quad \texttt{util/tikz\_geometry.py:506}.

\noindent\textbf{Soft spots.}\quad The check does not require $d \neq a$ (a degenerate ray); a coincident-with-vertex $d$ would cause a divide-by-zero caught by the helper and return \texttt{None} (silent skip).
The check also does not require $d$ to be inside $\angle bac$ --- it accepts $d$ on the \emph{external} bisector (perpendicular to the internal bisector at $a$), which gives equal angles to both legs but is on the opposite side.
Pair with \propname{same\_side} (IR-side) for prompts that require the internal bisector specifically.

\subsubsection{\propname{intersects}}
\label{app:pred:intersects}

\noindent\textbf{Signature.}\quad \texttt{intersects([[a, b], [c, d], p])}.

\noindent\textbf{Semantics.}\quad Lines $\overleftrightarrow{ab}$ and $\overleftrightarrow{cd}$ intersect at point $p$.

\noindent\textbf{Decision rule.}\quad Compute the line-line intersection $q$ via the standard determinant formula; return \false if the lines are parallel; otherwise return $\max(|q_x - p_x|, |q_y - p_y|) \leq \tau$.

\noindent\textbf{Reference.}\quad \texttt{util/tikz\_geometry.py:517} (intersection helper at \texttt{:119}).

\noindent\textbf{Soft spots.}\quad None observed for the line-line case; the parallel-line return-\false is correct (no intersection exists).

\subsubsection{\propname{equidistant\_from\_sides}}
\label{app:pred:equidistant_from_sides}

\noindent\textbf{Signature.}\quad \texttt{equidistant\_from\_sides([p, a, b, c])}.

\noindent\textbf{Semantics.}\quad Point \texttt{p} is equidistant from the three side-lines of $\triangle abc$ (i.e., \texttt{p} is the incentre or one of the three excentres).

\noindent\textbf{Decision rule.}\quad Compute the perpendicular distances $d_{ab}, d_{bc}, d_{ca}$ from $p$ to each side-line; return $|d_{ab} - d_{bc}| \leq \tau$ AND $|d_{bc} - d_{ca}| \leq \tau$.

\noindent\textbf{Reference.}\quad \texttt{util/tikz\_geometry.py:541}.

\noindent\textbf{Soft spots.}\quad Distance is to the side-\emph{line} (extended), not the side \emph{segment}; the three excentres are equidistant from the three lines and pass this check (\cref{ssec:ver_defn}).
Templates that mean ``incentre'' specifically should pair with three \propname{same\_side} co-checks ensuring \texttt{p} is on the same side of each side-line as the opposite vertex.

\subsection{Source-string predicates (decided against the TikZ source)}

\subsubsection{\propname{label\_present}}
\label{app:pred:label_present}

\noindent\textbf{Signature.}\quad \texttt{label\_present(["name"])}.

\noindent\textbf{Semantics.}\quad A label primitive for the named point appears in the TikZ source.

\noindent\textbf{Decision rule.}\quad Scan extracted labels (\texttt{tikz\_extraction.extract\_labels}) for \texttt{label\_point} and \texttt{label\_points} entries whose target name matches via \texttt{point\_names\_match} (case-insensitive, ignoring \LaTeX-math escapes).

\noindent\textbf{Reference.}\quad \texttt{util/tikz\_geometry.py:529}.

\noindent\textbf{Soft spots.}\quad This is a TikZ-source check, not a render check: a label primitive that exists in the source but is positioned outside the canvas, hidden by overlap, or otherwise visually invisible \emph{still passes}.
This is intentional: the predicate verifies the model's intent, not the renderer's layout.
Most scenarios use the dedicated \texttt{required\_labels} field (graded by \texttt{validate\_required\_labels}) rather than the symbolic-predicate form, which is why \propname{label\_present} appears infrequently in scenarios released as part of v1.
Both forms decide the same condition.

\subsubsection{\propname{mark\_present}}
\label{app:pred:mark_present}

\noindent\textbf{Signature.}\quad \texttt{mark\_present(["kind", "vertex"])} where \texttt{"kind"} is one of \texttt{"right\_angle"}, \texttt{"angle"}, \texttt{"length"}, \texttt{"tick"}.

\noindent\textbf{Semantics.}\quad A mark primitive of the requested kind, attached to the named vertex, appears in the TikZ source.

\noindent\textbf{Decision rule.}\quad Scan extracted marks (\texttt{tikz\_extraction.extract\_marks}) for entries whose \texttt{type} matches \texttt{kind} and whose \texttt{vertex}, \texttt{from}, or \texttt{to} field name-matches the vertex argument.

\noindent\textbf{Reference.}\quad \texttt{util/tikz\_geometry.py:535}.

\noindent\textbf{Soft spots.}\quad As with \propname{label\_present}, this is a source-string check: a mark primitive that would render off-canvas or that uses a non-standard TikZ command not recognised by the extractor will fail to register.
Per \cref{ssec:ver_defn}, no observed pilot record fails because a mark was rendered via raw \texttt{\textbackslash draw} primitives instead of the \texttt{\textbackslash tkzMark*} family the extractor recognises; the planned render-aware variant (parsing the rendered SVG for the actual right-angle glyph or tick stroke) is deferred to v1.1.

\subsection{Reserved predicates (declared but not yet wired into the eval verifier)}
\label{app:reserved_predicates}

Three predicates remain declared but unwired in v1 because, to a first approximation, all v1 scenarios that need them can be expressed without them.
They split into two distinct categories.

\paragraph{IR-side checker only: \propname{opposite\_side}, \propname{same\_side}.}
These are full first-class predicates in the IR layer (\texttt{ir/ir.py:662,671}; \texttt{ir/checks.py:154,169}), used by the \structured strategy during \emph{self}-validation before TikZ emission, but are not currently re-evaluated by the eval-side verifier (\texttt{util/tikz\_geometry.py}) on the rendered output.
\textbf{Signature.}\quad \texttt{opposite\_side([p, q, line\_a, line\_b])} and similarly \texttt{same\_side}.
\textbf{Semantics.}\quad The signed cross product $\vec{(\text{line\_a})(\text{line\_b})} \times \vec{(\text{line\_a})\,p}$ has the opposite (resp.\ same) sign as the corresponding quantity for $q$.
\textbf{Status.}\quad Not currently emitted by any v1 scenario, so the missing eval-side wiring has no observed effect on headline numbers.
Designed to pair with \propname{angle\_bisector}, \propname{equidistant\_from\_sides}, and other predicates whose adversarial soft spots are closed by adding an orientation co-check (\cref{ssec:ver_defn}).

\paragraph{Reserved for v1.1: \propname{not\_between}.}
\textbf{Signature.}\quad \texttt{not\_between([d, b, c])}.
\textbf{Semantics.}\quad The negation of \propname{point\_on\_segment}: $d$ may lie on line $\overleftrightarrow{bc}$ but is \emph{not} on segment $\overline{bc}$.
\textbf{Status.}\quad Implemented in neither layer and emitted by no v1 scenario; reserved as a v1.1 negative-incidence primitive for templates whose construction places a point on a line extension (e.g., the foot of an altitude that falls outside the triangle's base segment).

\subsection{Self-referential checks}
\label{ssec:self_referential_checks}

A predicate is \emph{self-referential} for a given strategy when the LLM's natural construction of the relevant geometric object makes the predicate's decision rule pass by symbolic identity.
The check still runs --- it can fail if the extractor mis-resolves coordinates or if the LLM chose a wrong construction primitive --- but it cannot fail because the LLM \emph{computed an answer incorrectly}, because the LLM did not compute an answer at all.

We observed three flavours of self-reference in v1, and the property language was designed to be transparent about each.

\paragraph{IR-derived points.} The \structured strategy emits IR primitives such as \texttt{point\_midpoint(M, A, B)}, \texttt{point\_centroid(G, A, B, C)}, \texttt{point\_intersection(P, line\_1, line\_2)}, and \texttt{point\_foot(H, X, line)}.
The compiler (\texttt{ir/to\_sympy.py}) evaluates these to numerical coordinates exactly per the textbook formulas; the predicate-language checks (\propname{midpoint}, \propname{centroid}, \propname{intersects}, \propname{point\_on\_line}) are the same formulas.
For this strategy, the predicate over an IR-derived point is \emph{tautological}: it is satisfied iff the extractor recovers the coordinate the compiler emitted.

\paragraph{TikZ-macro construction in \rawcode.} Capable LLMs writing raw TikZ delegate geometric arithmetic to \texttt{tkz-euclide} macros: $\backslash$\texttt{tkzDefMidPoint(A,B)}, $\backslash$\texttt{tkzInterLL(line\_1)(line\_2)}, $\backslash$\texttt{tkzCircumCenter(A,B,C)}, $\backslash$\texttt{tkzDefPointBy[rotation=...]}, etc.
These compute the same derived points as the IR but inside the LaTeX engine.
Because the predicate language and the macro library encode the same Euclidean theorems, predicates over points constructed this way are also tautological.
Empirically, in the headline run only $5$ of $\sim 600$ \texttt{tpl-t3-cen} records across all cells fail the \propname{centroid} predicate itself; the rest produce a $G$ within $10^{-9}$ of $(A+B+C)/3$ (\rawcode records using $\backslash$\texttt{tkzInterLL} on two medians satisfy this exactly by the median-concurrency theorem).

\paragraph{Literal-coordinate construction.} Only when the LLM hard-codes numerical coordinates directly (e.g., $\backslash$\texttt{tkzDefPoint(2.5,1.7)\{G\}} for a triangle with vertices summing to $(7.5, 5.1)$) is the predicate non-trivially testing arithmetic.
We did not observe this pattern for derived points in the v1 pilot.

\paragraph{Why retain self-referential checks.}
Three reasons.
(i)~The check is not vacuous: it catches extractor bugs, mis-named labels, and incorrect construction-primitive choices (e.g., emitting \texttt{point\_intersection} of two non-medians and labelling the result $G$).
(ii)~Removing them would push the headline metric toward strategies that hard-code coordinates over strategies that delegate construction, which inverts the methodological preference of \GeoGen for symbolic constructions.
(iii)~For diagram-generation \emph{benchmarks}, the LLM's choice of construction primitive (``the centroid is the intersection of any two medians'') is itself a meaningful skill; the predicate verifies that the LLM picked a primitive whose semantics agree with the prompt's intent.

\paragraph{Implication for headline interpretation.}
When most cells of a per-template column report identical-and-high strict-pass rates, the proximate cause is often that the predicate is self-referential and every observed strategy chose a satisfying construction.
The \texttt{tpl-t3-cen} case study (\cref{sssec:cen_case}) is the worked example: across $\sim 600$ headline-run records, only $5$ records fail \propname{centroid} itself, while $\sim 30$ fail on \emph{other} (non-self-referential) checks in the same scenario --- principally \propname{required\_labels} and timeouts.
A v1.1 hardening direction is to define \emph{construction-rendered} variants of self-referential predicates --- e.g., a centroid check that requires the three medians to be \emph{drawn} as detectable strokes in the SVG, in addition to $G$ being computationally correct --- which would shift the test from ``which primitive did the LLM use'' toward ``did the LLM make the construction visible to a viewer''.
We document this as future work (\cref{sec:discussion}) rather than v1 functionality.

\subsection{Tolerance and the verdict ladder}

All symbolic predicates use the same default tolerance $\tau = 5\times 10^{-3}$ in unit-coordinate space.
A scenario YAML may override this per-check via the \texttt{tol} field; in v1, $13/801$ scenarios use a per-check tolerance, all loosenings (e.g., \texttt{tol: 0.05}) for predicates over coordinates produced by trigonometric formulas where TikZ's printf-style \texttt{cos(35.0)} introduces a $\sim$1\% rounding step.
We chose absolute (not relative) tolerance because every \GeoGenBench scenario keeps coordinates in $[-10, 10]$ and a single uniform threshold simplifies adversarial analysis (\cref{ssec:ver_defn}).
For \GeoGen instantiations spanning wider coordinate ranges (architectural drawings, astronomical diagrams), a relative tolerance regime is the obvious extension.

A predicate's verdict is one of:
\begin{itemize}
\item \texttt{True} --- decision rule passed within $\tau$;
\item \texttt{False} --- decision rule failed;
\item \texttt{None} --- a required input coordinate was not in $\mathbf{X}$ (extractor failure or unimplemented derived-point construction), or, for \propname{label\_present}/\propname{mark\_present}, the TikZ source was not provided.
\end{itemize}
A \texttt{None} verdict downgrades the scenario from \texttt{pass} to \texttt{soft\_pass}; only \texttt{False} causes \texttt{fail}.
The \cref{ssec:ver_defn} pass closed $23$ \texttt{None}-verdicts in the pilot ($4.6\%$ of $498$ records, in two waves: $13$ from coordinate-extraction hardening and $10$ from implementing \propname{centroid}); the three reserved predicates above are the remaining known source of \texttt{None} verdicts in \GeoGenBench v1.
