\section{Reproduction instructions}
\label{app:reproduction}

\subsection{Pinned versions}
\label{app:pinned_versions}

All results in this paper were produced against repository commit \texttt{db17ae5} on \texttt{master} (with the paper-only commits in \texttt{feat/geogen-paper} layered on top). The pilot dataset in \cref{sec:results} was scored by the verifier at that commit; we re-graded all 498 pilot records against the same commit's verification logic and confirmed zero \texttt{gate\_status} drift versus the as-collected JSONL records.

The renderer Docker image is pinned to the tag \texttt{tikz-renderer:db17ae5}, built from the \texttt{renderer/} directory at the same commit. The renderer was migrated mid-development from a DVI pipeline to a PDF+\texttt{fontspec} pipeline (master commit \texttt{a74970f}); reproductions before that commit will produce visually different SVGs but should yield identical strict-pass verdicts because the verifier operates on the SymPy-compiled coordinate set extracted from the TikZ source, not on rendered SVG bytes.

\subsection{Quickstart}

Clone the repository, install the Python deps with \texttt{uv sync}, start the renderer container:
\begin{lstlisting}
git clone https://github.com/<anon>/geogen
cd geogen
git checkout db17ae5             # paper-pinned commit
uv sync
docker run -d -p 8001:8001 \
    ghcr.io/<anon>/tikz-renderer:db17ae5
\end{lstlisting}

\subsection{Reproducing the headline numbers}

The headline run in \cref{sec:results} can be reproduced with:
\begin{lstlisting}
bash evals/leaderboard_launch.sh
\end{lstlisting}
This iterates over (model $\times$ strategy) combinations against \texttt{evals/scenarios\_generated.yaml}, writing per-combo JSONL files into \texttt{evals/results/headline\_run/}.
Subsequent analysis:
\begin{lstlisting}
uv run python -m evals.leaderboard_analyze \
    --input-dir evals/results/headline_run/unified \
    --output-md REPORT.md --output-csv leaderboard.csv

uv run python -m evals.leaderboard_plot \
    --input-dir evals/results/headline_run/unified \
    --output-dir docs/figures/geogen-pilot
\end{lstlisting}

\subsection{Re-rolling the dataset}

To generate a contamination-free re-roll of the templated split with a different seed:
\begin{lstlisting}
uv run python -m evals.generate_scenarios \
    --output evals/scenarios_generated_v2.yaml \
    --seed 42
\end{lstlisting}
The procedural variant generator preserves the 30-template / 200-per-tier structure and is the recommended defence against suspected contamination.

\subsection{Re-grading without LLM calls}

Because the headline-run JSONL files store every model's TikZ output verbatim, the verifier can be re-run against a different commit's check semantics without spending another dollar on LLM tokens:
\begin{lstlisting}
uv run python -m evals.regrade \
    --input-dir evals/results/leaderboard_pilot_v2 \
    --output-dir evals/results/leaderboard_pilot_v3
\end{lstlisting}
We used this exact invocation to materialise \texttt{leaderboard\_pilot\_v3} from \texttt{leaderboard\_pilot\_v2} after pass~3 of \cref{ssec:ver_defn} (the \propname{centroid} predicate): 10/498 records promoted, 0 demotions.
The same pattern audited verdict stability across the renderer/IR migration described in \cref{app:pinned_versions} (498 records, zero gate-status changes pre-pass-3), and is also the right tool for any future hardening pass: re-grade first, re-run only if the verdicts shift in a way the new code cannot reproduce.

\subsection{Docker images and dataset card}

We release:
\begin{itemize}
\item \texttt{ghcr.io/<anon>/tikz-renderer:db17ae5} --- the TikZ$\to$SVG renderer (LuaLaTeX + \texttt{fontspec} + \texttt{tkz-euclide}, pinned to the paper commit). A floating \texttt{:latest} tag tracks the current \texttt{master} for active development; reviewers should pull the pinned tag.
\item \texttt{ghcr.io/<anon>/geogen-eval:db17ae5} --- the full Python evaluator with all dependencies pre-installed.
\item \texttt{datasets/geogen-v1.0} on Hugging Face --- the 801-scenario benchmark in JSONL (\texttt{bench\_generated.jsonl} with 600 templated records and \texttt{bench\_curriculum.jsonl} with 201 curriculum records), generated deterministically from the YAML sources in \texttt{benchmark/definitions/} by \texttt{benchmark/definitions/build\_release\_jsonl.py}.
\end{itemize}

\subsection{Croissant metadata and datasheet}

A \textbf{Croissant 1.0} metadata file accompanies the release at \texttt{benchmark/definitions/croissant.json}.
It conforms to \texttt{http://mlcommons.org/croissant/1.0}, declares both released JSONL files with their SHA-256 checksums, and exposes per-record \texttt{id}, \texttt{prompt}, \texttt{tier}, and \texttt{tags} fields for both the \texttt{templated} and \texttt{curriculum} splits.
The file additionally carries the Croissant Responsible-AI vocabulary fields covering data collection, annotation, preprocessing, biases, limitations, recommended uses, and social impact (mined from the Gebru et al.\ datasheet, \texttt{docs/datasheet.md}, and the discussion sections of this paper).
Validation passes against the reference \texttt{mlcroissant} Python implementation v1.1.0:
\begin{lstlisting}
.venv/bin/python -c "
import mlcroissant as mlc
ds = mlc.Dataset(jsonld='benchmark/definitions/croissant.json')
print(ds.metadata.ctx.issues.report() or 'OK')"
\end{lstlisting}
The complete \GeoGen datasheet (\citealp{gebru2021datasheets} format) is at \texttt{docs/datasheet.md}; the Hugging-Face-ready dataset card is at \texttt{benchmark/definitions/README.md} and uses the same content.

\subsection{Curriculum source attribution and licensing}
\label{app:curriculum_attribution}

The 201 curriculum-extracted scenarios (\cref{ssec:sources}) were prompted by lesson-level topical structure drawn from \textit{Bluebonnet Learning} K--12 mathematics, the open educational curriculum produced by Carnegie Learning, Inc.\ for the Texas Education Agency under the Texas State Board of Education's Instructional Materials and Technology Allotment process \citep{bluebonnet2024}.
Specifically, we used the Teacher's Editions of the Grade 6, Grade 7, Grade 8, and high-school Geometry volumes.
The Bluebonnet Learning materials are made publicly available by the Texas Education Agency at \url{https://learn.texas.gov/} for use by Texas public schools and the broader education community.

\paragraph{What we use from Bluebonnet, and what we release.}
Our pipeline uses Bluebonnet only as a \emph{topical scaffold}: we extract the lesson-level structure (chapter titles, lesson topics, and the geometric concepts each lesson exercises), and we prompt an LLM (\sonnet) to author a fresh, student-style diagram-drawing request that exercises that concept.
We do not redistribute, paraphrase, or otherwise include any portion of the Bluebonnet textbook prose, figures, problem statements, solutions, or pedagogical commentary in the released benchmark.
The released artefacts are: (a) our LLM-generated diagram-drawing prompts (each is a short, self-contained natural-language request with no excerpted Bluebonnet text); (b) the corresponding property-list rubrics we authored against those prompts; and (c) the per-scenario lineage tags identifying which Bluebonnet chapter and lesson \emph{topic} a scenario was inspired by, so that downstream users can replicate or extend the topical coverage.
The Bluebonnet attribution lives in each curriculum scenario's \texttt{metadata.lineage} field in the released JSONL.

\paragraph{Licensing.}
The Bluebonnet Learning curriculum carries its own publication terms set by the Texas Education Agency and Carnegie Learning, Inc.\ for the Bluebonnet programme; we do not assert any rights in those source materials, and any use of the Bluebonnet curriculum itself must be made under those terms.
Our derivative artefacts --- the LLM-generated prompt rewrites, the property-list rubrics, the verifier code, the renderer, the Croissant metadata, the datasheet, and this paper --- are released under the MIT License.
Users wishing to access the underlying Bluebonnet Learning materials should obtain them directly from the Texas Education Agency Bluebonnet portal (\url{https://learn.texas.gov/}); we do not redistribute any Bluebonnet content.
