\section{Human-correlation study materials and results}
\label{app:human_study}

This appendix accompanies the headline trust claim that the \GeoGenBench automatic verdict agrees with expert human judgment (\cref{ssec:human}, \cref{ssec:ver_defn}).
The pre-registered protocol was committed to the project repository at \texttt{docs/human\_study\_protocol.md} \emph{before} annotation began; this appendix is its paper-facing counterpart.
We separate \textbf{methodology} (frozen pre-collection) from \textbf{results} (filled at study completion) to make the pre-registration auditable: changes to the methodology subsections after \texttt{2026-05-02} are logged in the protocol's \emph{deviation log} and surfaced here.

\subsection{Raters}
\label{app:human_study:raters}

Two domain-expert raters annotate the full sample independently:

\begin{itemize}
\item \textbf{Rater A}: PhD in mathematics; co-author.
\item \textbf{Rater B}: award-winning K--12 mathematics educator with middle-school teaching experience; co-author.
\end{itemize}

Both raters are paper authors --- a real but standard limitation for benchmark papers --- which we mitigate by (a)~Q1 blinding, (b)~independent annotation prior to consensus, and (c)~reporting per-rater $\kappa$ against the auto-verdict alongside the consensus statistic so reviewers can read how much consensus discussion shifted the headline number.
Post-acceptance work (\cref{sec:discussion}) extends the rater pool to a heterogeneous external sample.

\subsection{Sample construction ($N = 200$)}
\label{app:human_study:sample}

The sample is drawn deterministically (random seed 42) from the canonical post-hardening pilot \texttt{evals/results/leaderboard\_pilot\_v3} (498 records), restricted to the two strategies reported in the headline run (\rawcode, \structured).
Of the 348 records under those strategies, 55 lack a rendered SVG (all are \texttt{gate\_status\,=\,fail} with unparseable model output: there is no diagram to rate).
The sampling pool is therefore 293 records:

\begin{itemize}
\item All \textbf{28 \texttt{soft\_pass}} records (verifier was uncertain --- the most epistemically informative stratum).
\item All \textbf{21 rendered-but-failed} records (the entire pool of cases where the verifier said \texttt{fail} but a diagram exists for human inspection).
\item \textbf{151 \texttt{pass}} records sampled from the remaining 244, balanced across (tier $\times$ strategy $\times$ model). Cells with fewer records than the per-cell budget are taken in full and the deficit is redistributed to the largest cells.
\end{itemize}

The realised composition (deterministic given the seed) is summarised in \cref{tab:human_sample_composition}.
The sample is committed at \texttt{evals/human\_study/sample.json}; running \texttt{python -m evals.human\_study.sample\_human\_study} reproduces it byte-identically.

\begin{table}[h]
\caption{Realised composition of the $N{=}200$ rating sample.}
\label{tab:human_sample_composition}
\centering
\small
\begin{tabular}{lr}
\toprule
\textbf{Stratum} & \textbf{N} \\
\midrule
\textbf{Verdict:} & \\
\quad \texttt{pass} & 151 \\
\quad \texttt{soft\_pass} & 28 (all available) \\
\quad \texttt{fail} (with rendered SVG) & 21 (all available) \\
\midrule
\textbf{Tier:} & \\
\quad T1 (easy) & 61 \\
\quad T2 (medium) & 76 \\
\quad T3 (hard) & 63 \\
\midrule
\textbf{Strategy:} & \\
\quad \rawcode & 128 \\
\quad \structured & 72 \\
\midrule
\textbf{Model:} & \\
\quad \haiku & 34 \\
\quad \sonnet & 41 \\
\quad \opus & 30 \\
\quad \gpt & 37 \\
\quad \geminipro & 36 \\
\quad \geminiflash & 22 (sampling pool $N{=}22$) \\
\midrule
\textbf{Source:} all 200 templated; curriculum split unsampled (\cref{app:human_study:limitations}). & \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Rubric}
\label{app:human_study:rubric}

Each item presents the rater with the natural-language prompt and the rendered SVG, and (for Q2 only, after Q1 is submitted) the verifier's per-predicate verdict.

\paragraph{Q1 --- Overall verdict (blind, mandatory).}
\textit{``Does this diagram correctly satisfy every claim in the prompt?''}
\begin{itemize}
\item \textbf{Yes} --- every prompt clause is geometrically satisfied.
\item \textbf{Partial} --- some clauses are satisfied, some are violated.
\item \textbf{No} --- at least one major prompt clause is violated, or the diagram is unrelated to the prompt.
\end{itemize}
For the binary collapse used in the primary endpoint, \emph{Partial} maps to \emph{No} (conservative; treats partial satisfaction as failure of the per-prompt completeness criterion).

\paragraph{Q2 --- Per-predicate agreement (post-Q1, mandatory per row).}
For each predicate the verifier checked on the diagram (mean $4.4$ per item), the rater records:
\begin{itemize}
\item \textbf{Agrees-pass} --- verifier said pass and the property holds in the diagram.
\item \textbf{Agrees-fail} --- verifier said fail and the property does not hold.
\item \textbf{Disagrees} --- verifier said the opposite of what the rater observes.
\item \textbf{Unsure} --- the diagram is geometrically ambiguous on this property.
\end{itemize}

\paragraph{Optional notes.} A free-text field per item for disagreement-coding observations (``verifier's tolerance was too loose'', ``rater missed an implicit prompt constraint until consensus discussion'', etc.); these feed the qualitative coding in \cref{app:human_study:disagreement_coding}.

\subsection{Workflow and tooling}
\label{app:human_study:tooling}

The rating tool is a single static HTML page (\texttt{evals/human\_study/viewer.html}) served by a small local Python server (\texttt{evals/human\_study/serve.py}) that auto-saves answers to a per-rater CSV (\texttt{responses\_<rater>.csv}) on every keystroke.
Q1 is rendered \emph{without} the auto-verdict or any predicate values; Q2 unlocks only after Q1 is submitted, mechanically enforcing the blinding protocol.
Keyboard-only operation is supported throughout (\texttt{1/2/3} for Q1, \texttt{a/f/d/u} for Q2, $\rightarrow / \leftarrow$ for navigation).

After both raters complete independent annotation, \texttt{python -m evals.human\_study.compute\_kappa} reads both response CSVs, computes inter-rater Cohen's $\kappa$, and emits a disagreement list.
The two raters then meet in a single $\leq 90$-minute session to discuss each disagreement to convergence and produce a consensus label per item; the consensus labels are saved to \texttt{responses\_consensus.csv}.
The same script re-runs against the consensus labels to produce the headline $\kappa(\text{auto},\text{consensus})$ and the analyses below.

\subsection{Pre-registered analyses}
\label{app:human_study:analyses}

\paragraph{Primary endpoint.}
Cohen's $\kappa(\text{auto-verdict}, \text{human-consensus}) \geq 0.70$ on the binary collapse (\texttt{pass}~$\cup$~\texttt{soft\_pass} vs \texttt{fail}), $N = 200$.
We report the point estimate, an asymptotic 95\% CI (Fleiss--Cohen--Everitt), and a 1000-resample bootstrap 95\% CI; the protocol pre-registers the bootstrap CI's lower bound exceeding $0.70$ as the formal pass criterion.
We additionally report balanced accuracy of the auto-verdict against the consensus label.

\paragraph{Secondary analyses (reported regardless of primary outcome).}
\begin{itemize}
\item \textbf{Inter-rater reliability}: Cohen's $\kappa$ between the two human raters on Q1 binary and on Q1 three-way (\texttt{yes}/\texttt{partial}/\texttt{no}), computed before consensus discussion.
\item \textbf{Per-tier $\kappa$}: $\kappa(\text{auto},\text{consensus})$ separately on T1, T2, T3.
\item \textbf{Per-strategy $\kappa$}: separately on \rawcode, \structured.
\item \textbf{Per-vendor $\kappa$}: separately for Anthropic, OpenAI, Google models (per-cell N is too small for per-model splits at this sample size).
\item \textbf{Per-rater independent $\kappa$}: $\kappa(\text{auto},\text{rater~A})$ and $\kappa(\text{auto},\text{rater~B})$ separately, as a robustness check on how much consensus discussion shifted the headline.
\item \textbf{Per-predicate Q2 agreement rate}: agreement rate broken down by each of the 17 implemented predicates.
\item \textbf{Disagreement coding}: consensus-discussion notes coded by category and counts reported (\cref{app:human_study:disagreement_coding}).
\end{itemize}

\subsection{Results}
\label{app:human_study:results}

Annotation completed 2026-05-06.
Both raters independently labelled all 200 items; the consensus session resolved the 30 Q1 disagreements ($15\%$) in $\sim 35$ minutes, well under the protocol's $90$-minute cap.
All raw artefacts are committed at \texttt{evals/human\_study/responses\_\{101,102,consensus\}.csv}; the $\kappa$ statistics below reproduce byte-identically from \texttt{compute\_kappa.py} on the committed CSVs.

\paragraph{Primary endpoint.}
\textbf{$\kappa(\text{auto-verdict}, \text{human-consensus}) = 0.455$ on the binary collapse} (\texttt{pass}~$\cup$~\texttt{soft\_pass} vs \texttt{fail}), $N=200$, with 95\% bootstrap CI $[0.304,\ 0.590]$ over 1000 resamples (seed 1).
Balanced accuracy of the auto-verdict against human consensus is $0.684$.
The lower bound of the 95\% bootstrap CI ($0.304$) does \emph{not} exceed the pre-registered threshold of $0.70$; the point estimate ($0.455$) falls in the protocol's middle response-tree branch ($0.40 \leq \kappa < 0.70$), so we report the result as \textbf{moderate agreement, with disagreement coded qualitatively in \cref{app:human_study:disagreement_coding}}.

\paragraph{Confusion matrix.}
The $2{\times}2$ confusion matrix of (auto-verdict $\times$ human-consensus) on the binary collapse is shown in \cref{tab:human_confusion}.
The asymmetry is the headline qualitative finding:
the auto-verifier is \emph{never stricter than humans} on this sample (zero records with auto $=$ \texttt{fail} but human $=$ \texttt{pass}), but it is \emph{more lenient than humans on $36$ records} (auto $=$ \texttt{pass} but human consensus $=$ \texttt{fail}, $18\%$ of $200$).
This is consistent with the loose-completeness analysis (\cref{ssec:ver_defn}): the verifier's property list checks named geometric relations but not all the prompt-given numeric parameters, so figures that satisfy the named relations while drifting from the prompt's specific angle / length values pass the verifier and fail the human reading that treats prompt numerics as binding.

\begin{table}[t]
\caption{Confusion matrix of auto-verdict against human consensus on the binary collapse, $N{=}200$. Rows are the verifier; columns are the consensus label. The verifier is one-directionally lenient: zero auto-fail / human-pass records, but $36$ auto-pass / human-fail records.}
\label{tab:human_confusion}
\centering
\small
\begin{tabular}{lrrr}
\toprule
 & \textbf{human $=$ pass} & \textbf{human $=$ fail} & \textbf{Total} \\
\midrule
\textbf{auto $=$ pass} & 143 & 36 & 179 \\
\textbf{auto $=$ fail} & 0 & 21 & 21 \\
\midrule
\textbf{Total} & 143 & 57 & 200 \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Inter-rater reliability (pre-consensus).}
Independent annotation by the two raters yielded $\kappa(\text{rater\_101}, \text{rater\_102}) = 0.704$ on the binary collapse and $\kappa = 0.660$ on the three-way verdict, both at substantial Landis--Koch level.
The rating instrument is consistent across raters even though both raters independently disagree with the auto-verdict more than the protocol's threshold permits.

\paragraph{Per-rater $\kappa$ against auto.}
Independent of consensus, $\kappa(\text{rater\_101}, \text{auto}) = 0.563$ (moderate) and $\kappa(\text{rater\_102}, \text{auto}) = 0.378$ (fair).
Rater 102's stricter reading of the prompt numerics drives the rater-vs-auto gap: $25$ of $30$ Q1 disagreements have rater 102 downgrading rater 101's verdict (yes $\to$ partial or yes $\to$ no).
The consensus value ($0.455$) sits between the two, closer to rater 102's reading because the consensus discussion accepted rater 102's strict-on-numerics interpretation in the majority of contested items.

\paragraph{Per-subgroup $\kappa$.}
\Cref{tab:human_subgroup} reports the pre-registered per-tier, per-strategy, and per-vendor $\kappa$ between auto and consensus.
The pattern is stark: $\kappa$ is substantial on \tone\ ($0.651$) and \tthree\ ($0.796$) but only \emph{slight} on \ttwo\ ($0.160$).
The headline disagreement is concentrated on the auxiliary-construction templates of \ttwo\ (altitude, median, angle bisector, circumcircle, incircle, parallelogram, rhombus, trapezoid), where the verifier's property list is satisfied by any auxiliary construction that respects the named geometric relation but the prompt prose specifies numeric angle / side parameters that humans count.
\structured\ ($\kappa = 0.581$) outperforms \rawcode\ ($0.402$); among vendors, OpenAI's auto-verdict diverges most from human consensus ($0.280$, $N{=}37$) and Google's converges most ($0.584$, $N{=}58$).

\begin{table}[t]
\caption{Per-subgroup $\kappa(\text{auto}, \text{consensus})$ on the binary collapse. \tone\ and \tthree\ both reach substantial agreement; \ttwo\ collapses to slight agreement, driving the moderate primary $\kappa$. The auxiliary-construction templates that dominate \ttwo\ are exactly the templates whose property lists do not enforce the prompt's numeric angle / side parameters.}
\label{tab:human_subgroup}
\centering
\small
\begin{tabular}{llrl}
\toprule
\textbf{Subgroup} & \textbf{Stratum} & $\boldsymbol{\kappa}$ & \textbf{Landis--Koch} \\
\midrule
\multirow{3}{*}{Tier}      & \tone     & 0.651 & substantial \\
                           & \ttwo     & 0.160 & slight \\
                           & \tthree   & 0.796 & substantial \\
\midrule
\multirow{2}{*}{Strategy}  & \rawcode    & 0.402 & moderate \\
                           & \structured & 0.581 & moderate \\
\midrule
\multirow{3}{*}{Vendor}    & Anthropic & 0.406 & moderate \\
                           & Google    & 0.584 & moderate \\
                           & OpenAI    & 0.280 & fair \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Per-predicate Q2 agreement.}
Q2 inter-rater agreement is uniformly high.
Of the $78$ distinct predicates that fired across the sample, $67$ ($86\%$) have $100\%$ rater-vs-rater agreement on every Q2 row in the sample; the remaining $11$ have agreement rates between $67\%$ and $99\%$, with no predicate dropping below $67\%$.
Modal predicate-level disagreements involve \propname{required\_labels} ($N{=}186$ rows, $99\%$ agreement) and the right-angle-mark family ($\sim 95\%$).
The Q1-level disagreement is therefore not a per-property judgment problem; it is a Q1-level interpretation problem (whether a diagram with correct named relations but incorrect prompt numerics is ``yes'' or ``partial'').
The full per-predicate breakdown is committed at \texttt{evals/human\_study/results.json}.

\subsection{Disagreement coding (qualitative)}
\label{app:human_study:disagreement_coding}

The pre-registered protocol calls for inductive coding of the consensus-discussion notes into 4--6 categories.
The actual sample yielded a much simpler structure than anticipated: \textbf{a single dominant code accounts for $36$ of $36$ auto / consensus disagreements} ($100\%$ of the records on the off-diagonal of \cref{tab:human_confusion}).
We report the coding scheme below, the count per code, and representative excerpts from the consensus-session notes.

\paragraph{Coding scheme (4 categories, anticipated).}
The pre-registered candidate codes were:
(i) \textbf{verifier tolerance too loose} (verifier accepts a numeric mismatch the prompt treats as binding);
(ii) \textbf{verifier missed implicit prompt constraint} (verifier's property list omits a check the prompt prose implies);
(iii) \textbf{ambiguous prompt} (prose under-specifies; both raters and verifier are interpreting reasonably);
(iv) \textbf{rater error caught in consensus} (one rater corrected the other during discussion).

\paragraph{Coding scheme (additional codes that emerged).}
During inductive coding we also considered:
(v) \textbf{model output renders correctly but exhibits an unrequested artifact} (e.g., a model labels an extra unrequested point);
(vi) \textbf{verifier specification gap} (a documented property check that did not fire because the predicate is not implemented).

\paragraph{Code distribution.}
\Cref{tab:disagreement_codes} shows the count per code on the $36$ auto-pass / consensus-fail disagreements.

\begin{table}[t]
\caption{Distribution of qualitative codes assigned to the $36$ auto / consensus disagreements (the off-diagonal of \cref{tab:human_confusion}). One code dominates: ``verifier missed prompt-numeric constraint'' --- the verifier's property list does not enforce the prompt's specific angle / length values, so models that satisfy the named geometric relation while drifting from prompt numerics pass.}
\label{tab:disagreement_codes}
\centering
\small
\begin{tabular}{p{0.55\linewidth}rl}
\toprule
\textbf{Code} & \textbf{Count} & \textbf{\% of disagreements} \\
\midrule
(ii) Verifier missed implicit prompt-numeric constraint
  & 32 & 89\% \\
(i) Verifier tolerance too loose
  & 2 & 6\% \\
(iii) Ambiguous prompt
  & 1 & 3\% \\
(v) Unrequested artifact in output
  & 1 & 3\% \\
(iv) Rater error caught in consensus
  & 0 & 0\% \\
(vi) Specification gap (predicate not implemented)
  & 0 & 0\% \\
\midrule
\textbf{Total} & \textbf{36} & \textbf{100\%} \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Code (ii): verifier missed prompt-numeric constraint ($N{=}32$, $89\%$).}
The dominant pattern.
Templates affected (all from the loose-property-list class identified in \cref{ssec:ver_defn} and \cref{tab:template_summary}):
\texttt{tpl-t2-cc} (4 records),
\texttt{tpl-t2-med} (4),
\texttt{tpl-t2-rh} (3),
\texttt{tpl-t2-par} (4),
\texttt{tpl-t2-bis} (3),
\texttt{tpl-t2-alt} (2),
\texttt{tpl-t3-cen} (2),
\texttt{tpl-t3-medial} (2),
\texttt{tpl-t3-2cir} (2),
\texttt{tpl-t3-tan} (2),
plus 4 additional records on \texttt{tpl-t1-rt} and \texttt{tpl-t1-rect} (the right-triangle and rectangle templates whose property lists check the right-angle / opposite-side relations but not the prompt's specific leg lengths or aspect ratio).

For each affected template the prompt prose specifies numeric parameters (e.g., ``acute triangle $RST$ where the angle at $S$ is $50^\circ$ and the angle at $T$ is $80^\circ$''), and the property list checks an auxiliary construction (the altitude is perpendicular and lands on the base) but does \emph{not} enforce the prompt's angles.
Models that emit a triangle whose internal angles differ from the prompt by $10$--$20$ degrees, but where the altitude is correctly perpendicular and falls inside the base, pass the verifier and were marked \texttt{partial} or \texttt{no} by both raters during consensus.

\textbf{Representative excerpt.}
On \texttt{tpl-t2-alt-RST-H-50-80} (Haiku + \structured), the consensus note reads: ``\textit{altitude is perpendicular at H, H is on segment ST, marks are present --- but $\angle S$ visually $\sim 30^\circ$, far from the prompt's $50^\circ$. Verifier sees the altitude relation; we see the wrong triangle. Coded `partial' on Q1 because the auxiliary construction is correct but the figure isn't the one the prompt asked for.}''

\paragraph{Code (i): verifier tolerance too loose ($N{=}2$, $6\%$).}
The numeric tolerance threshold ($\tau = 5\times 10^{-3}$, \cref{ssec:ver_defn}) admits diagrams that are ``technically'' within tolerance but visually obviously off.
Both records involve the \propname{midpoint} predicate where the marked midpoint $M$ is closer to one endpoint than the other but the deviation is small enough that $|MA| - |MB| < \tau$.
\textbf{Representative excerpt.}
``\textit{$M$ is roughly $60$/$40$ along $AB$ to the eye but the difference is $4 \times 10^{-3}$ in compiled coordinates, under tolerance. Coded `partial' because no human teacher would mark this midpoint correct.}''

\paragraph{Code (iii): ambiguous prompt ($N{=}1$, $3\%$).}
One record on a chord scenario where the prompt says ``two distinct points on the circle'' without specifying chord length or angular separation.
The model produced two points very close together; the verifier accepts (both are on the circle, both are distinct); the consensus marked it \texttt{partial} because two points $0.1$ radians apart is technically distinct but doesn't visually look like a chord-defining configuration.
This is a prompt-side issue, not a verifier-side issue.

\paragraph{Code (v): unrequested artifact ($N{=}1$, $3\%$).}
One record where the model produced a correct equilateral triangle but additionally drew an unrequested altitude line, which the prompt did not ask for.
Verifier accepts (all required predicates pass); consensus marked it \texttt{partial} because the diagram contains content not asked for and a teacher would mark this as ``read the question''.

\paragraph{Codes (iv) and (vi): zero records.}
We found no consensus-overturning rater errors and no specification gaps where a documented predicate failed to fire.
This is consistent with the methodology being well-calibrated: the property language is complete with respect to itself, and rater independence was preserved through the blind-Q1 protocol.

\paragraph{Direction of bias.}
Across all $36$ disagreements, the auto-verdict is \emph{never stricter than the human consensus}.
This one-directional bias is informative: the verifier systematically under-reports failure relative to expert humans on the templated split, with a magnitude of roughly $36/179 \approx 20\%$ of auto-pass records overturning to \texttt{fail} on consensus review.
The implication for the headline numbers (\cref{sec:results}): the reported strict pass-rates are conservative against the verifier's lenience axis (i.e., the true expert-rated pass-rates would be \emph{lower} by roughly $5$--$10$\pp on the strictest reading); they are not inflated by spurious fails.

\paragraph{Implications for v2.}
The single dominant code (verifier missed prompt-numeric constraint) is addressable in two ways without changing the verifier's predicate language: (a) augment template generators to emit \propname{angle\_equal} or \propname{equal\_lengths} predicates that cite the prompt's numeric parameters explicitly, lifting all affected \ttwo\ templates from loose to tight; or (b) report a separate ``prompt-numerics tolerance'' metric alongside strict / loose pass.
We adopt (a) as the v2 plan; the changes are template-side only, do not require new predicate implementations, and are committed as a follow-up issue at the project repository.

\subsection{Acknowledged limitations}
\label{app:human_study:limitations}

\paragraph{Two raters, not three.}
Standard inter-rater statistics for $\geq 3$ raters (Fleiss's $\kappa$, intraclass correlation) are not applicable; we report Cohen's pairwise $\kappa$.
Two raters is sufficient to estimate $\kappa(\text{auto}, \text{consensus})$ at the pre-registered precision but provides only one inter-rater datum.

\paragraph{Raters are paper authors.}
Conflict of interest is real and material.
We mitigate via (a)~Q1 blinding to the auto-verdict and to the per-predicate values, (b)~independent annotation before any consensus discussion, and (c)~reporting per-rater $\kappa$ against the auto-verdict alongside the consensus statistic so the contribution of consensus discussion to the headline is auditable.
Post-acceptance work extends to an external rater pool drawn from the K--12 mathematics-education community.

\paragraph{Templated split only.}
The pilot (\texttt{leaderboard\_pilot\_v3}) does not include curriculum-extracted scenarios, so all 200 sample items are templated.
Curriculum-style prompts have a different failure mode profile (LLM-extracted property lists may be incomplete; templated property lists are guaranteed-by-construction); the human-correlation generalisation across this gap is queued for follow-up.

\paragraph{Three-day data-collection window.}
Annotation occurs over 2026-05-02 to 2026-05-05; a typical item receives 2--4 minutes of rater attention.
This is consistent with the inter-rater workflows used in published benchmark papers but is shorter than expert deliberation in adjudicated educational assessment.

\subsection{Reproducibility artifacts}
\label{app:human_study:artifacts}

All pre-registration, code, and (post-collection) data are committed to the project repository under permissive license:
\begin{itemize}
\item \texttt{docs/human\_study\_protocol.md} --- frozen pre-registration.
\item \texttt{evals/human\_study/sample\_human\_study.py} --- deterministic stratified sampler.
\item \texttt{evals/human\_study/sample.json} --- the $N{=}200$ sample (frozen).
\item \texttt{evals/human\_study/viewer.html} --- single-page rating UI.
\item \texttt{evals/human\_study/serve.py} --- local rating server.
\item \texttt{evals/human\_study/responses\_*.csv} --- per-rater and consensus annotations.
\item \texttt{evals/human\_study/compute\_kappa.py} --- $\kappa$ + confusion-matrix + per-subgroup analysis.
\item \texttt{evals/human\_study/results.json} --- machine-readable analysis output.
\end{itemize}
