\section{Introduction}
\label{sec:intro}

Modern frontier LLMs are being asked to do things that were science-fiction five years ago, including authoring entire visual artifacts: slide decks, worksheets, technical illustrations, and engineering diagrams.
The most-deployed visual artifact in the K--12 educational market is the \emph{geometry diagram}: a 2D plane figure paired with a problem statement.
Yet every existing geometry benchmark we are aware of evaluates the inverse direction --- given a diagram, answer a question about it --- and provides no way to evaluate whether a model can \emph{produce} a correct diagram from a textual description.
This is the gap \GeoGenBench fills.

\subsection{Motivating example}

\begin{quote}\itshape
``Draw an acute triangle $EFG$ with vertex angles $60^\circ$, $70^\circ$ at $E$ and $F$. Construct the altitude from $G$ meeting $EF$ at $H$.''
\end{quote}
A correct rendering must satisfy: (a) $\angle E = 60^\circ$, $\angle F = 70^\circ$, (b) $H$ lies on segment $EF$ (not its extension), (c) $\angle GHE = 90^\circ$, (d) the triangle is acute.
\GeoGenBench encodes each clause as a machine-checkable predicate against the SymPy-compiled coordinates of the model's rendered TikZ output (\cref{fig:teaser}).
On this scenario (\texttt{tpl-t2-alt-EFG-H-60-70} in our pilot), \haiku and \opus pass under both strategies; \gpt with \rawcode places $H$ on the line $EF$ but \emph{outside} the segment (extension); \sonnet with \rawcode draws an obtuse triangle.
The verifier flags each failure with a specific predicate verdict, not a holistic score, enabling actionable per-check error analysis.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    promptbox/.style={draw, rounded corners, fill=blue!6, align=left,
                      text width=39mm, minimum height=34mm,
                      inner sep=5pt, font=\small\itshape},
    diagrambox/.style={draw, rounded corners, fill=gray!4, align=center,
                       minimum width=42mm, minimum height=34mm, inner sep=4pt},
    verdictbox/.style={draw, rounded corners, fill=green!6, align=left,
                       text width=42mm, minimum height=34mm,
                       inner sep=5pt, font=\scriptsize\ttfamily},
    role/.style={font=\scriptsize\itshape, color=gray!55!black, align=center},
    arrlbl/.style={font=\tiny\itshape, color=gray!50!black},
    >=Stealth, line width=0.5pt, node distance=8mm
]
  %--- header roles row ---
  \node[role] (lbl1) at (0, 0.6) {prompt};
  \node[role] (lbl2) at (5.6, 0.6) {model output (rendered)};
  \node[role] (lbl3) at (11.2, 0.6) {symbolic predicate verdicts};

  %--- prompt (left) ---
  \node[promptbox, anchor=north] (prompt) at (lbl1 |- 0,0)
    {``Draw an acute triangle $EFG$ with vertex angles $60^{\circ}$, $70^{\circ}$ at $E$ and $F$.
      Construct the altitude from $G$ meeting $EF$ at $H$.''};

  %--- rendered diagram (middle) — drawn in TikZ ---
  % Coordinates: E=(0,0), F=(5,0). Angle at E = 60°, angle at F = 70° → G ≈ (3.07, 5.31). H = foot ≈ (3.07, 0).
  \node[diagrambox, anchor=north] (dia) at (lbl2 |- 0,0) {};
  \begin{scope}[shift={($(dia.south) + (-1.45,0.4)$)}, scale=0.5]
    \coordinate (E) at (0,0);
    \coordinate (F) at (5,0);
    \coordinate (G) at (3.07,5.31);
    \coordinate (H) at (3.07,0);
    \draw[thick] (E) -- (F) -- (G) -- cycle;
    \draw[densely dashed, thick] (G) -- (H);
    % right-angle marker at H (small square opening up-left toward G)
    \draw[thick] (H) ++(-0.42,0) -- ++(0,0.42) -- ++(0.42,0);
    % vertex labels
    \node[anchor=north east, font=\small] at (E) {$E$};
    \node[anchor=north west, font=\small] at (F) {$F$};
    \node[anchor=south,      font=\small] at (G) {$G$};
    \node[anchor=north,      font=\small] at ($(H)+(0.15,-0.05)$) {$H$};
    % angle arcs + labels
    \draw[thin] (0.7,0) arc (0:60:0.7);
    \draw[thin] (4.3,0) arc (180:110:0.7);
    \node[font=\scriptsize] at (1.05,0.34) {$60^{\circ}$};
    \node[font=\scriptsize] at (3.95,0.34) {$70^{\circ}$};
  \end{scope}

  %--- verdicts (right) ---
  \node[verdictbox, anchor=north] (verdict) at (lbl3 |- 0,0)
    {\textcolor{green!50!black}{\textbf{\checkmark}}\ angle\_equal($\angle E,60^{\circ}$)\\[1pt]
     \textcolor{green!50!black}{\textbf{\checkmark}}\ angle\_equal($\angle F,70^{\circ}$)\\[1pt]
     \textcolor{green!50!black}{\textbf{\checkmark}}\ right\_angle($G,H,E$)\\[1pt]
     \textcolor{green!50!black}{\textbf{\checkmark}}\ point\_on\_segment\\[-2pt]
     \hphantom{\textbf{\checkmark}\ \ }($H,E,F$)\\[1pt]
     \textcolor{green!50!black}{\textbf{\checkmark}}\ label\_present\\[-2pt]
     \hphantom{\textbf{\checkmark}\ \ }($\{E,F,G,H\}$)\\[3pt]
     \rmfamily\textbf{verdict:} \textcolor{green!50!black}{\textbf{strict\_pass}}};

  %--- arrows between the three boxes (above the boxes' midline) ---
  \draw[->, thick] (prompt.east)  -- (dia.west)
        node[arrlbl, midway, above] {LLM emits}
        node[arrlbl, midway, below] {TikZ$\to$SVG};
  \draw[->, thick] (dia.east)     -- (verdict.west)
        node[arrlbl, midway, above] {compile to}
        node[arrlbl, midway, below] {SymPy + check};
\end{tikzpicture}
\caption{\textbf{A worked \GeoGenBench scenario, end-to-end.}
The model receives only the natural-language prompt (left).
It emits TikZ that we render to SVG and re-compile to symbolic geometry (centre).
The per-scenario property list (right) is decided in microseconds against the SymPy coordinates: every clause becomes a typed predicate with a \texttt{true/false} verdict, no human grading, no LLM-as-judge.
A scenario passes only when \emph{all} predicates are satisfied; failures localise to the specific predicate that broke.}
\label{fig:teaser}
\end{figure}

\subsection{Methodology and instance}
\label{ssec:naming}

This paper contributes both a \emph{methodology} for building automatically-verifiable benchmarks of generative geometric reasoning and a \emph{first instance} of that methodology.
We use \textbf{\GeoGen} for the methodology --- two coupled ideas, formalised in \cref{ssec:methodology}: (i) procedural co-generation of prompts and property-list rubrics from shared source code, and (ii) a render-then-symbolic-verify pipeline driven by a declarative predicate language --- and \textbf{\GeoGenBench} for the 801-prompt plane-geometry benchmark in which we instantiate it.
The distinction matters: future \GeoGen-family benchmarks for coordinate-plane algebra, 3D solid geometry, transformation-group constructions, and engineering or architectural drawings would inherit the methodology and the predicate-language design (\cref{app:property_reference}) while extending the per-domain symbolic primitives (\cref{ssec:future_methodology}).

\subsection{Contributions}

\begin{enumerate}
\item \textbf{The \GeoGen methodology and reusable verification toolkit.}
We formalise a two-pillar methodology for generative-geometric-reasoning benchmarks: (a) \emph{procedural co-generation}, in which a small Python generator emits the natural-language prompt and the machine-checkable rubric from the same source code, eliminating drift between the question and its grading criteria and providing a built-in contamination defence (reviewers can re-roll the engine with new seeds); and (b) \emph{render-then-symbolic-verify}, in which the model's text-emitted output is compiled to symbolic objects against which a versioned, declarative predicate language (\cref{ssec:properties}, \cref{app:property_reference}) is decided in milliseconds.
We release both pillars as reusable code (template engine + verifier + 17-predicate language v1 + a soundness audit, \cref{sec:hardening}, that closes 4.6\% of silently-skipped pilot checks across three hardening passes).
\item \textbf{The \GeoGenBench instance: 801 prompts, six frontier models, and a leaderboard with two deployment-relevant findings plus a cross-vendor portability finding.}
\GeoGenBench is the methodology's first instantiation: 600 procedurally co-generated prompts (released as \texttt{bench\_generated.yaml}) plus 201 LLM-extracted curriculum prompts for breadth (released as \texttt{bench\_curriculum.yaml}).
We benchmark six frontier models from three vendors (\opus, \sonnet, \haiku, \gpt, \geminipro, \geminiflash) under two generation strategies (raw TikZ; structured intermediate-representation JSON with check-and-revise), reporting --- at the \cref{sec:hardening}-hardened verifier on a $600$-scenario $\times$ $3$-repeat draw --- a strict-pass spread of \textbf{$58$\pp} ($35.9$\% to $94.1$\%) across the eleven measurable cells, a \textbf{$\mathbf{7{\times}}$ cost-with-equivalent-pass spread} between \gpt + \structured ($94.1$\% at \dollars{0.075}/scen) and \opus + \structured ($93.3$\% at \dollars{0.532}/scen), and a vendor-specific deployment-portability finding (\cref{ssec:gemini_finding}): \geminiflash's schema-state cap blocks structured generation entirely.
The dataset, the template engine, the verification toolkit, the leaderboard data, and a Docker-packaged TikZ$\to$SVG renderer are all released under permissive licenses (\cref{app:reproduction}).
\end{enumerate}

\subsection{Headline empirical results}

Two findings from the post-hardening headline run, in order of importance:

\begin{itemize}
\item \textbf{Strategy effect is uneven across models.}
The \structured strategy lifts every model on strict pass, but the gain ranges from $+2.2$ to $+24.7$\pp:
\gpt ($91.9\% \to 94.1\%$, $+2.2$\pp), \haiku ($70.5\% \to 75.3\%$, $+4.8$\pp), \sonnet ($81.2\% \to 93.4\%$, $+12.2$\pp), \opus ($71.6\% \to 93.3\%$, $+21.7$\pp), \geminipro ($58.4\% \to 83.1\%$, $+24.7$\pp).
Models that already do well at raw TikZ (\gpt) gain little from scaffolding; models that struggle at raw (\opus, \geminipro) gain the most.
The pre-hardening verifier reported gaps of $10$--$30$\pp; the \cref{sec:hardening} audit attributes about half of the early signal to silent skips of \texttt{\textbackslash pgfmathsetmacro}-style coordinates that \rawcode outputs use more often than IR-compiled outputs.
\item \textbf{Frontier convergence: three flagship models tie within $1$\pp at IR strategy, but cost differs by $7\times$.}
Under \structured, \gpt ($94.1$\% strict, \dollars{0.075}/scenario), \sonnet ($93.4$\%, \dollars{0.088}), and \opus ($93.3$\%, \dollars{0.532}) span only $0.8$\pp of strict pass at a cost ratio of $7.1\times$; \gpt and \sonnet are essentially tied while \opus carries no measurable quality premium for $6\times$ the cost.
\opus + \rawcode (\dollars{0.262}/scenario, $71.6$\%) is the most decisively dominated cell: \gpt + \rawcode (\dollars{0.058}, $91.9$\%) gives $20$\pp better pass at $4.5\times$ lower cost.
The Pareto frontier collapses to two cells (cheapest and best-quality), with all three frontier+IR cells essentially co-located.
\end{itemize}

\subsection{Roadmap}

\Cref{sec:related} situates \GeoGenBench against geometry-understanding benchmarks, math benchmarks without diagrams, and image-generation benchmarks.
\Cref{sec:design} describes the dataset and the property language.
\Cref{sec:verification} audits verifier soundness and completeness, including a quantified false-positive rate from adversarial probing and the \cref{sec:hardening} that lifted 4.6\% of pilot records from soft-pass to strict-pass.
\Cref{sec:strategies} describes the two generation strategies.
\Cref{sec:results} presents the leaderboard, Pareto analysis, per-tier and per-template breakdowns, and failure-mode taxonomy.
\Cref{sec:discussion} discusses what \GeoGenBench does and does not measure, limitations, and threats to validity.
