\section{Methodology and instance}
\label{sec:design}

\subsection{The \GeoGen methodology}
\label{ssec:methodology}

We separate \emph{methodology} from \emph{instance} from the outset.
\GeoGen is a methodology for building automatically-verifiable benchmarks of generative geometric reasoning; \GeoGenBench (\cref{ssec:task}~onward) is its first instantiation, for plane Euclidean geometry.
The methodology rests on two pillars:

\paragraph{Pillar 1: Procedural co-generation of prompts and rubrics.}
For domains where the construction admits a Python expression --- which is essentially every domain in classical geometry --- both the natural-language prompt and the machine-checkable property list can be emitted from the same source code.
A single \texttt{altitude\_template($A$, $B$, $C$, $H$)} function, for example, returns both a prose request to draw the altitude and the predicate list \texttt{[right\_angle($A$, $H$, $B$), point\_on\_segment($H$, $B$, $C$)]} that any correct rendering must satisfy.
This eliminates the drift problem that haunts hand-authored or LLM-extracted benchmarks (where the grading rubric and the question can disagree silently) and provides a strong contamination defence (reviewers can re-roll the engine with new seeds, label sets, or numeric parameters and obtain a fresh split with the same difficulty distribution).
For domains where procedural co-generation is impractical --- 3D constructions, multi-step proof figures, or scenarios requiring extra-mathematical context --- the methodology supports a fallback mode (\cref{ssec:sources}): hand-authored or LLM-extracted prompts paired with hand- or LLM-authored property lists.
This trades the correctness-by-construction guarantee for breadth, and we report metrics on the two modes separately.

\paragraph{Pillar 2: Render-then-symbolic-verify with a declarative predicate language.}
The model emits source code in a text-emitting renderer's input language (TikZ here, but also SVG, Asymptote, OpenSCAD, or any other text-format vector medium); a containerised renderer compiles that to its visual output; an extractor compiles the same source to a coordinate-bearing symbolic representation (SymPy here); a fixed-vocabulary, typed \emph{predicate language} (\cref{ssec:properties}, formalised in \cref{app:property_reference}) is decided against the symbolic representation in microseconds per predicate.
Three properties make this work as a benchmark substrate.
First, the predicate language is \emph{declarative} --- prompt authors specify \emph{what must be true}, not \emph{how to check it} --- so adding a new construction never requires writing a new checker.
Second, the predicate vocabulary is \emph{versioned}: \cref{app:property_reference} formalises v1 (17 symbolic + 2 source-string predicates); future \GeoGen instantiations extend it (e.g.\ \texttt{face\_incidence} for 3D, \texttt{point\_on\_curve} for coordinate plane) without disturbing existing benchmarks.
Third, the recipe is \emph{renderer-agnostic}: the codebase ships an SVG-output strategy alongside TikZ to demonstrate that swapping the text-emitting target is a one-paragraph change in the strategy layer (\cref{ssec:strategies}).

\paragraph{Why these two pillars together.}
Each pillar partially addresses the canonical reviewer attack on a generative-reasoning benchmark.
Pillar~1 addresses ``how do we know your ground truth is right?'' (it is correct by construction wherever procedural generation applies, and the engine source is part of the release).
Pillar~2 addresses ``how do we know your verifier is sound?'' (declarative predicates have stable, auditable semantics; a soundness audit is reported in \cref{sec:hardening}; cross-version verdict stability is documented in \cref{app:pinned_versions}).
Together they let the leaderboard numbers in \cref{sec:results} be interrogated end-to-end: every prompt traces to a Python generator, every check traces to a SymPy decision procedure, every disagreement is reproducible without an LLM in the loop.
The remainder of this section instantiates both pillars for plane Euclidean geometry as \GeoGenBench.

\subsection{Task formulation}
\label{ssec:task}

A \GeoGenBench scenario is a tuple
\[
\mathcal{S} = \langle \text{prompt},\ \mathcal{P},\ \mathcal{L},\ \mathcal{C} \rangle,
\]
where the \textbf{prompt} is a natural-language request to draw a geometric figure (e.g., \emph{``Draw an acute triangle $ABC$ with the altitude from $A$ meeting $BC$ at $H$. Label all four points and mark the right angle at $H$.''}), $\mathcal{P}$ is a list of \emph{expected geometric properties} the rendered figure must satisfy, $\mathcal{L}$ is a set of \emph{required labels} (point names that must appear in the output), and $\mathcal{C}$ is an optional set of \emph{structural checks} (e.g., ``must contain at least one quadrilateral'').

A model must produce a TikZ source string that compiles to a valid SVG.
The system then (1)~\textbf{renders} the TikZ via a containerised LuaLaTeX + dvisvgm pipeline; (2)~\textbf{verifies} by compiling the diagram's coordinate set and applying each property in $\mathcal{P}$ against a SymPy symbol table, comparing $\mathcal{L}$ against TikZ string content, and applying each structural check; and (3)~\textbf{scores} the scenario as one of \texttt{pass} (all checks passed), \texttt{soft\_pass} (no failures but at least one check skipped), \texttt{fail} ($\geq 1$ hard failure), or \texttt{gen\_failure}/\texttt{render\_failure} (no usable output).

Crucially, the model is \emph{not} given access to the property list during generation --- only the natural-language prompt.
Properties are the grading rubric, not the input.

\subsection{Scenario sources}
\label{ssec:sources}

\paragraph{Templated scenarios (provably correct ground truth).}
A procedural template engine emits scenarios in which the prompt prose and the expected-property list are produced \emph{from the same source code}, guaranteeing they are mutually consistent.
We use 30 templates grouped by tier (9~\tone, 10~\ttwo, 11~\tthree) covering right/equi\-lateral/iso\-sce\-les triangles, squares/rectangles/parallelo\-grams/rhombi/trapezoids, segment+midpoint, parallel/perpendicular pairs, circle+point, altitude/median/angle-bisector/circumcircle/incircle/perpendicular-bisector/chord constructions, right-triangle altitude to hypotenuse, medial triangle, centroid+medians, parallel lines + transversal, isosceles altitude bisects base, tangent to circle, cyclic quadrilateral, two intersecting circles, Thales' theorem, triangle with two altitudes, and inscribed angle in a semicircle.
Per-template variation axes (10 vertex-triple label sets, 6 quad-label sets, 8 pair-label sets, numeric parameter sweeps such as Pythagorean leg pairs, and 7 prompt-opener variants) multiplied by round-robin per-tier capping yield \textbf{600 scenarios} (200 per tier).
Property correctness is by construction: the same Python function emits both the prose and the property list.
The released artefact is \texttt{benchmark/definitions/bench\_generated.yaml}.

\paragraph{Curriculum-extracted scenarios (broader topic coverage).}
A separate pipeline (\texttt{curriculum/to\_benchmark.py}) extracts curriculum structure from K--12 geometry textbooks and uses an LLM to convert each into a student-style diagram request, plus structural and property-level expected outputs.
This adds \textbf{201 scenarios} ($[36, 128, 37]$ across tiers) covering topics templated scenarios do not (3D solids, transformations, coordinate-overlay constructions, multi-step proofs requiring auxiliary lines), at the cost of LLM-authored ground truth that is spot-checkable but not guaranteed correct in the same way templated GT is.
We report metrics decomposed by source so reviewers can trust the templated half as a calibration anchor while the curriculum half stress-tests breadth.
The released artefact is \texttt{benchmark/definitions/bench\_curriculum.yaml}.

\subsection{Property language}
\label{ssec:properties}

Each scenario carries a list of expected geometric properties drawn from a fixed vocabulary of \textbf{14 symbolic predicate types implemented in the eval verifier} plus 2 source-string check types, all enforced against a coordinate symbol table extracted from the rendered TikZ.
A further 4 predicates (\propname{centroid}, \propname{opposite\_side}, \propname{same\_side}, \propname{not\_between}) are declared in the IR-side checker but currently silent-skipped by the eval verifier; we document them transparently in \cref{app:reserved_predicates} and account for their effect on the headline numbers in \cref{sec:hardening}.
Each predicate is decided in microseconds with a numeric tolerance $\tau = 5\times 10^{-3}$ (justified empirically against human ratings; see \cref{sec:verification}).
\Cref{tab:predicates} summarises the language; \cref{app:property_reference} gives the full signature, semantics, decision rule, and code reference for each predicate.

\begin{table}[t]
\caption{The \GeoGen predicate language v1. Symbolic predicates are exact checks against compiled coordinates; source-string checks are intentionally weaker (they verify a TikZ primitive that \emph{would} render the mark appears in the source) and contribute only to \emph{loose} pass. Four additional predicates declared in the IR layer are not yet wired into the eval verifier --- see \cref{app:reserved_predicates}.}
\label{tab:predicates}
\centering
\small
\begin{tabular}{lll}
\toprule
\textbf{Predicate} & \textbf{Args} & \textbf{Decision} \\
\midrule
\multicolumn{3}{l}{\emph{Symbolic (exact, implemented in v1)}} \\
\propname{right\_angle} & $[a, o, b]$ & $|\vec{oa}\cdot\vec{ob}| < \tau$ \\
\propname{midpoint} & $[m, a, b]$ & $\max(|m_x-(a_x{+}b_x)/2|, |m_y-(a_y{+}b_y)/2|) < \tau$ \\
\propname{collinear} & $[p_1, p_2, p_3]$ & $|\vec{p_1 p_2}\times\vec{p_1 p_3}| < \tau$ \\
\propname{equal\_lengths} & $[[a,b], [c,d], \dots]$ & all pairwise $\big||AB| - |CD|\big| < \tau$ \\
\propname{parallel} & $[[a,b], [c,d]]$ & $|\vec{ab}\times\vec{cd}| < \tau$ \\
\propname{perpendicular} & $[[a,b], [c,d]]$ & $|\vec{ab}\cdot\vec{cd}| < \tau$ \\
\propname{point\_on\_line} & $[p, a, b]$ & $|\vec{ab}\times\vec{ap}| < \tau$ \\
\propname{point\_on\_segment} & $[p, a, b]$ & on line + $\big||ap|+|pb|-|ab|\big| < \tau$ \\
\propname{point\_on\_circle} & $[p, o, r]$ & $\big||OP|-|OR|\big| < \tau$ \\
\propname{tangent} & $[[L_1,L_2], O, T]$ & $T$ on line $L_1L_2$ + $\vec{OT} \perp \vec{L_1L_2}$ \\
\propname{angle\_bisector} & $[d, a, b, c]$ & $|\angle bad - \angle dac| < \tau$ \\
\propname{angle\_equal} & $[[a_1, b_1, c_1], [a_2, b_2, c_2]]$ & $|\angle a_1 b_1 c_1 - \angle a_2 b_2 c_2| < \tau$ \\
\propname{intersects} & $[[a,b], [c,d], p]$ & line-line intersection $\approx p$ \\
\propname{equidistant\_from\_sides} & $[p, a, b, c]$ & three side-line distances equal $< \tau$ \\
\midrule
\multicolumn{3}{l}{\emph{Source-string (soft)}} \\
\propname{label\_present} & $[\text{name}]$ & label primitive for \texttt{name} appears in TikZ source \\
\propname{mark\_present} & $[\text{kind}, \text{vertex}]$ & mark primitive of \texttt{kind} at \texttt{vertex} appears in source \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Verification pipeline}
\label{ssec:pipeline}

The pipeline is illustrated in \cref{fig:pipeline}.
The two-renderer choice (TikZ $\to$ SVG via Docker by default; pure-Python SVG renderer as fallback) decouples the benchmark from any specific TeX install while staying reproducible.

\begin{figure}[t]
\centering
\begin{tikzpicture}[node distance=4mm, font=\small,
    box/.style={draw, rounded corners, minimum width=22mm, minimum height=8mm, align=center}]
  \node[box] (prompt) {prompt};
  \node[box, right=of prompt] (llm) {LLM strategy\\ \scriptsize{\rawcode / \structured / \recipe}};
  \node[box, right=of llm] (tikz) {TikZ source};
  \node[box, below=of tikz] (latex) {LuaLaTeX\\(container)};
  \node[box, left=of latex] (svg) {SVG};
  \node[box, left=of svg] (sympy) {SymPy compile\\ + property check};
  \node[box, below=of svg] (verdict) {gate\_status \\ $\in \{$pass, soft\_pass, fail$\}$};
  \draw[->] (prompt) -- (llm);
  \draw[->] (llm) -- (tikz);
  \draw[->] (tikz) -- (latex);
  \draw[->] (latex) -- (svg);
  \draw[->] (svg) -- (sympy);
  \draw[->] (sympy) -- (verdict);
  \draw[->] (svg) -- (verdict);
\end{tikzpicture}
\caption{The \GeoGenBench verification pipeline. The LLM produces TikZ; the TikZ is rendered to SVG by a containerised LuaLaTeX/dvisvgm; the SVG and the compiled SymPy geometry are independently checked against the per-scenario property list; the result is a single \texttt{gate\_status}.}
\label{fig:pipeline}
\end{figure}

\paragraph{Gate-status semantics.}
We mark a scenario \texttt{fail} if any check failed; \texttt{soft\_pass} if no checks failed but at least one was skipped (typically rendering-only); \texttt{pass} if all checks ran and returned true.
Headline tables use \emph{strict-pass} (\texttt{pass} only); soft-pass acts as a diagnostic for ``did the model get the geometry right but skip a label?'' versus ``did it produce something illegible.''

\subsection{Tiered difficulty}
\label{ssec:tiers}

Templated scenarios are tier-labelled at template-definition time:
\begin{itemize}
\item \textbf{\tone (easy)}: single named shape, no auxiliary construction.
\item \textbf{\ttwo (medium)}: one shape + one auxiliary construction.
\item \textbf{\tthree (hard)}: multi-step or composed construction.
\end{itemize}
Curriculum scenarios inherit tier labels from the LLM curriculum extractor; we audit them but do not retier.

\subsection{Dataset statistics}
\label{ssec:stats}

\begin{table}[t]
\caption{\GeoGenBench dataset composition.}
\label{tab:dataset_stats}
\centering
\small
\begin{tabular}{lrrrrrr}
\toprule
\textbf{Split} & \textbf{\#scen.} & \textbf{\tone} & \textbf{\ttwo} & \textbf{\tthree} & \textbf{Avg.\ \#props} & \textbf{Constructions} \\
\midrule
Templated & 600 & 200 & 200 & 200 & 3.4 & 30 templates \\
Curriculum & 201 & 36 & 128 & 37 & TBD & 73 unique tags \\
\midrule
\textbf{Total} & \textbf{801} & 236 & 328 & 237 & --- & 100+ \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Reproducibility and contamination defence}
\label{ssec:repro}

The procedural template engine is part of the release; reviewers can re-roll the templated split with a different seed or new label sets to detect contamination.
We commit to releasing the templated set as \emph{train} and a held-back roll of the engine as \emph{test}.
All renders are deterministic: the scenario YAML schema and the property checker are pure functions of the input, with no nondeterminism in grading.

\subsection{Known limitations of the verification toolkit}
\label{ssec:design_limits}

We document soft spots up front rather than be surprised by them in review:
\begin{enumerate}
\item \propname{point\_on\_segment} originally checked collinearity but not segment bounds; this was a true source of false positives in pre-hardening pilot data and is fixed in \cref{sec:hardening}.
\item \propname{label\_present} and \propname{mark\_present} are TikZ-string checks; a model that emits \texttt{\textbackslash tkzMarkRightAngle} passes the check even if the mark would visually clip off-canvas.
These checks are intentionally classified as soft and contribute to \texttt{soft\_pass}, not \texttt{pass}.
\item \propname{tangent} verifies the line is at radial distance $r$ from the centre but does not separately verify that $T$ itself lies on the circle; we recommend pairing with \propname{point\_on\_circle} in templates.
\item No 3D / solid-geometry checks. Curriculum scenarios that mention cones or spheres exercise the renderer but cannot be symbolically verified by the current toolkit.
\item Tolerance $\tau = 5\times 10^{-3}$ is fixed and uniform across check types; per-type tolerances are left for future work.
\end{enumerate}
