\section{Models and strategies}
\label{sec:strategies}

\subsection{Models}
\label{ssec:models}

We evaluate six production frontier models that span $\sim$20$\times$ in price and represent three API families (Anthropic, OpenAI, Google):

\begin{itemize}
\item \opus (Anthropic): \dollars{15}/\$75 per million input/output tokens.
\item \sonnet (Anthropic): \dollars{3}/\$15.
\item \haiku (Anthropic): \dollars{1}/\$5.
\item \gpt (OpenAI Responses API): \dollars{5}/\$20.
\item \geminipro (Google AI Studio API, paid tier, $\leq$200K context): \dollars{1.25}/\$10.
\item \geminiflash (Google AI Studio API, paid tier, $\leq$200K context): \dollars{0.30}/\$2.50.
\end{itemize}

All runs use temperature $= 1.0$ (default for both APIs).
Costs are computed using each model's published per-token rate as of April 2026; \cref{sec:results} reports both raw token counts and dollarised costs.

\subsection{Strategies}
\label{ssec:strategies}

\GeoGenBench ships three reference strategies as comparators, each implementing a common \texttt{SubstanceStrategy} interface.
The benchmark itself does not depend on any of these --- any text-emitting model can be benchmarked by routing its TikZ output through the verification pipeline.
The strategies sit on a spectrum from ``maximum LLM freedom'' (\rawcode) to ``maximum scaffolding'' (\recipe); comparing them isolates the contribution of structured intermediate representations and retrieval.

\paragraph{\rawcode (single-shot direct TikZ).}
The LLM is given the prompt plus a TikZ tutorial system prompt and produces TikZ via a \texttt{render\_diagram(tikz: str)} tool call.
It may call the tool multiple times to revise based on the rendered output, but no symbolic checks are run.
This is the zero-scaffolding baseline.

\paragraph{\structured (schema-constrained intermediate representation).}
The LLM is given a Pydantic schema for an intermediate \texttt{DiagramIR} representation (point declarations, segment declarations, intersections, circles, render ops, and \emph{check} declarations) and must emit IR JSON.
Code (not the LLM) compiles IR $\to$ SymPy objects $\to$ TikZ $\to$ SVG.
Validation against the IR's check list happens \emph{before} render, so the strategy can self-correct: if a declared check fails on the LLM's IR, the strategy revises (up to $N=3$ retries).
This is the strategy we recommend for production deployment.

\paragraph{\recipe (high-level DSL with few-shot retrieval).}
A cheaper ``selector'' model picks 0--N relevant \emph{recipes} (curated DSL examples like \texttt{circumcircle.yaml}, \texttt{altitude.yaml}) which are prepended as in-context examples for the main model.
The main model emits a \texttt{RecipeDSL} (high-level operations like \texttt{triangle}, \texttt{altitude}, \texttt{circumcircle}); a lowering pass compiles to \texttt{DiagramIR}, then the \structured pipeline runs.
This is the most heavily scaffolded strategy: the LLM picks from a constrained menu of pre-built constructions, trading creative freedom for built-in geometric correctness.

\paragraph{Beyond TikZ: an alternate text-emit target.}
The methodology is renderer-agnostic --- the verifier does not care whether the model emits TikZ, raw SVG, Asymptote, or any other text-emittable vector format, provided we can compile it back to a coordinate set.
The codebase ships \texttt{strategies/raw\_svg.py} (and a self-revising variant) which routes the model directly to SVG output, bypassing TikZ and the LuaLaTeX container entirely.
Because raw SVG and raw TikZ measure essentially the same thing --- ``can the model emit a text-format vector diagram with correct geometry'' --- we do not include \texttt{raw\_svg} in the headline leaderboard; it is intended for deployments where a TeX-free runtime is required, and is documented in the released codebase rather than benchmarked here.

\subsection{Cost and latency methodology}
\label{ssec:cost_methodology}

For each (model, strategy, scenario, repeat) we record (a)~total input + output tokens across all API calls (including any retries), (b)~wall-clock duration from prompt submission to rendered SVG, and (c)~the token-multiplied USD cost at published rates.
The \emph{mean cost per scenario} reported in \cref{sec:results} is the arithmetic mean of per-scenario costs across the pilot, including failed scenarios (a model that frequently retries before failing pays for the retries).
Per-call latency excludes time spent in our verification pipeline.

\subsection{Pilot configuration}
\label{ssec:pilot_config}

For the pilot reported in \cref{sec:results} we ran a stratified 30-scenario sample (10 per tier, drawn round-robin across 23 of 30 templates) with one repeat per (model, strategy, scenario) cell.
The full headline run uses the complete 600-scenario templated split with three repeats and reports per-cell standard error.
The \structured strategy permits up to three internal check-fail retries before recording final verdict.
We document the rate-limit ceiling we encountered with \opus + \structured at \cref{ssec:headline}.
