\section{Discussion}
\label{sec:discussion}

\subsection{What \GeoGenBench measures and what it does not}

\GeoGenBench measures whether a model can translate a textual specification of a 2D plane-geometry construction into a globally-consistent metric realisation that satisfies a stated set of relational properties.
This is a narrower task than ``geometric reasoning'' broadly construed --- it does not require proof, does not test 3D solid geometry, does not exercise coordinate-plane algebra, and does not measure \emph{teaching utility} (whether a diagram makes a concept clear).
We chose this scope because (i)~it is the part of geometric reasoning where automatic, exact, fast verification is achievable today; (ii)~it is the dominant deployed task for LLMs in K--12 worksheet authoring; and (iii)~it generalises in obvious ways (the same property language extends to coordinate-plane and 3D once the relevant SymPy primitives are wired in).

\subsection{Templated vs.\ curriculum: where they diverge}

Templated scenarios have provably-correct ground truth (prose and properties are emitted from the same source code) and bound the model's task tightly enough that we can quantify completeness rates per-template (\cref{ssec:completeness}).
Curriculum scenarios cover topics the templates do not (3D solids, transformations, coordinate-overlay constructions, multi-step proofs) but rely on LLM-authored ground truth that is spot-checkable but not guaranteed correct in the same way.
We report metrics decomposed by source so reviewers can trust the templated half as a calibration anchor while the curriculum half stress-tests breadth.
\TODO{Once full headline run completes, report the templated-vs-curriculum delta per-model so reviewers can see which split each finding survives on.}

\subsection{Limitations}

\begin{enumerate}
\item \textbf{2D Euclidean only.}
Curriculum scenarios that mention cones / spheres exercise the renderer but cannot be symbolically verified by the current toolkit.
\item \textbf{No proof construction.}
\GeoGenBench tests diagram production, not the propositional reasoning that justifies the construction.
A model that draws a correct diagram but cannot explain \emph{why} it is correct still passes.
\item \textbf{Single repeat in the pilot.}
The headline numbers in \cref{sec:results} are 1-repeat estimates; we have no run-to-run variance.
The full headline run uses 3 repeats; until then, the strict-pass differences below 5\pp should be read as suggestive only.
\item \textbf{Rate-limit ceiling on \opus + \structured.}
We report $N=12$ for that single cell (\cref{ssec:headline}); this is documented as a deployment finding rather than a noise issue.
\item \textbf{Tolerance is fixed.}
The $5\times 10^{-3}$ threshold was tuned for human-rating agreement on a small sample and is uniform across check types; per-type tolerances are left for future work.
\item \textbf{Loose-completeness is the modal case.}
$\sim 37\%$ of templated scenarios are loose (\cref{ssec:completeness}): the property list pins named relations but leaves continuous parameters free.
This is intentional and matches educational practice, but it means a passing verdict cannot be read as ``produced \emph{the} canonical diagram,'' only as ``produced \emph{a} diagram consistent with the prompt's stated constraints.''
\end{enumerate}

\subsection{Threats to validity}

\paragraph{Contamination.}
LLMs may have seen procedurally-similar geometry problems during training.
The procedural template engine ships with the benchmark, allowing reviewers to roll new contamination-free splits with a different seed; we further commit to releasing a held-out test split that is revealed only at evaluation time.
The curriculum split is naturalistic and harder to contamination-defend; we report templated-only metrics in the headline and curriculum metrics as a robustness check.

\paragraph{Soft-pass semantics.}
Pre-hardening, soft-pass made some configurations look better than they were.
\Cref{sec:hardening} converted 23/498 records (4.6\%) of pilot soft-passes to strict-pass over three hardening passes, and we report both numbers throughout.
We use \emph{strict pass} as the headline metric to avoid the criticism.

\paragraph{Verifier soundness.}
\Cref{ssec:adversarial} enumerated 5 predicates with measurable soft spots; all five are addressable by either hardening (\propname{point\_on\_segment}) or template-side mitigation (pairing with a complementary predicate).
We estimate $< 5\%$ of strict passes are vulnerable to an adversarial construction; the in-progress human-correlation study (\cref{ssec:human}, \cref{app:human_study}) will tighten this estimate at submission completeness.

\paragraph{LLM-judged check authoring (curriculum split).}
The curriculum-extracted scenarios use an LLM to translate textbook problems into property lists; the LLM may emit incorrect or incomplete property lists.
We spot-check this on a sample but cannot guarantee per-scenario correctness in the same way templated scenarios provide.
Headline metrics are reported templated-only; curriculum metrics are reported as a robustness check.

\subsection{Implications for deployment}

The pilot's two unambiguous Pareto-dominance findings (\cref{ssec:pareto}) have direct deployment implications:

\begin{itemize}
\item \textbf{For raw TikZ generation, the marginal cost of the frontier model buys nothing.}
\haiku + \rawcode dominates \opus + \rawcode at $19\times$ lower cost.
A team deploying an LLM-driven worksheet author should default to the cheap-tier model unless they have evidence their domain is in the Pareto-undominated region.
\item \textbf{At the mid-tier price point, model + strategy choice matters more than chasing the next model release.}
\sonnet + \recipe dominates \gpt + \structured at the same cost.
The optimal scaffolding is model-family-specific; we recommend benchmarking the strategy on the deployment target rather than assuming the leaderboard transfers.
\item \textbf{The recipe-DSL strategy is unstable below the frontier-model tier.}
\sonnet uses recipe well; \haiku and \gpt do not (worse than their \rawcode).
Custom DSLs with built-in geometric priors do not transfer; teams considering this approach should validate per-model.
\end{itemize}

\subsection{Extending the methodology}
\label{ssec:future_methodology}

Both pillars of \GeoGen (\cref{ssec:methodology}) are designed to extend beyond plane geometry.
Pillar~1 (procedural co-generation) is reusable wherever a domain admits a Python-expressible construction --- the engine in \texttt{evals/generate\_scenarios.py} is $\sim$1{,}000 lines and the per-template extension cost is $\sim$50 lines per construction; we expect the same template-engine pattern to apply to coordinate-plane algebra and most ruler-and-compass extensions of plane Euclidean geometry, with proportionally larger investments for 3D solids and transformations.
Pillar~2 (render-then-symbolic-verify) extends along the predicate-language axis: the v1 vocabulary in \cref{app:property_reference} covers plane-Euclidean relations; future \GeoGen-family benchmarks would publish v2, v3, \dots adding domain primitives without disturbing existing benchmarks (the predicate-language versioning pattern is intentional).
Two extensions are already in flight at the codebase level and could be released as separate \GeoGen-family benchmarks:
\begin{itemize}
\item \textbf{GenExam-Math integration.}
\citet{wang2025genexam} introduce a multidisciplinary text-to-exam-image dataset with LLM-judged scoring.
We have re-encoded a 91-prompt geometry-relevant subset (HIGH-relevance plus medium-cartesian) as a \GeoGen-style benchmark with declarative property checks (\texttt{benchmark/definitions/bench\_genexam.yaml}); a separate report will compare strict-pass rates between automatic verification and the GenExam vision-LM judge on the overlap.
This is an empirical test of whether vision-LM judges and SymPy-decided predicates agree on text-to-diagram tasks where both can be applied.
\item \textbf{Beyond plane geometry.}
The predicate language (\cref{app:property_reference}) extends to coordinate-plane algebra (point-on-curve, intersection-of-curves), 3D solids (face-incidence, dihedral angles), and transformation-group constructions (rotation/reflection/translation equivalence) by adding the relevant SymPy primitives.
Engineering and architectural drawings are similar: any domain where ``correctness'' decomposes into a finite list of relational properties decidable against a symbolic ground truth is a candidate for the same recipe.
The bottleneck is predicate-language design, not symbolic computation --- and the predicate language is the part future \GeoGen instantiations would inherit unchanged.
\end{itemize}
