\begin{abstract}
Existing geometry benchmarks measure whether language models can \emph{interpret} a given diagram and answer a question about it.
None measures whether a model can \emph{produce} a correct diagram from a description --- yet diagram production is increasingly the deployed task, as LLMs are asked to author worksheets, slides, and technical illustrations.
We introduce \textbf{\GeoGen}, a methodology for building automatically-verifiable benchmarks of generative geometric reasoning, and \textbf{\GeoGenBench}, its first instantiation.
\GeoGen pairs two ideas: (i)~\emph{procedural co-generation} of prompts and rubrics from shared source code, eliminating drift between the question and its grading criteria; and (ii)~a \emph{render-then-symbolic-verify} pipeline that compiles the model's text-emitted output (TikZ here) to symbolic objects and decides a list of declarative geometric predicates in milliseconds --- no human grading, no vision-language judge.
\GeoGenBench instantiates the methodology for plane Euclidean geometry: \textbf{801} natural-language prompts (e.g., ``draw an acute triangle $ABC$ with the altitude from $A$ meeting $BC$ at $H$'') paired with machine-checkable properties (``$\angle AHB$ is right; $H$ lies on segment $BC$'') drawn from a 17-predicate language.
Across \textbf{six frontier models from three vendors} (\opus, \sonnet, \haiku, \gpt, \geminipro, \geminiflash) and \textbf{three generation strategies} on a stratified pilot, we find that \emph{the cost--quality Pareto frontier is dominated by cheap-model configurations}: \haiku with direct TikZ generation reaches 83\% strict pass at \dollars{0.016}/scenario, beating \opus with the same strategy (80\%, \dollars{0.328}) on both axes --- 3\pp more correct diagrams at $21\times$ lower cost.
A generic intermediate-representation strategy adds a further 4--20\pp on top for vendors whose constrained-generation supports the IR's schema; we identify one vendor (Gemini Flash) where a hard schema-state cap blocks it entirely, and report this as a deployment-portability finding.
We also report a verification-reliability audit: \cref{sec:hardening} closes 13 of 315 (4.1\%) silently-skipped checks in the pilot via two extractor improvements, and we report both pre- and post-audit numbers throughout for transparency.
We release the dataset, the procedural template engine that generated it (allowing reviewers to roll new contamination-free splits), the verification toolkit, and a public leaderboard.
\end{abstract}
